\bf \Large A Model Evaluation Study for Treatment Planning of Laser Induced Thermal Therapy

### 2015-04-16 Reviewer_response.tex

\documentclass[letterpaper,12pt]{report}
\usepackage{fixltx2e}
\usepackage{amssymb,amsfonts,amsmath}
\usepackage{color,graphicx}
\usepackage{siunitx}
\usepackage{fixltx2e}
\usepackage[nomarkers,figuresonly]{endfloat}
\usepackage[left=0.7in,right=0.7in,top=0.7in,bottom=0.7in]{geometry}
\newcommand{\numberofpatientsLOOCV}{22 }
\begin{document}
Answer to reviewers\\
\\
Reviewer 1:\\
Comments to the Author\\
This is a well-written paper on an evaluation study of
two mathematical models. However, there are several issues with the current version as follow:\\

\begin{enumerate}
\item
The authors stated, in the abstract, that the objective
was to evaluate the trade-off between the
time investment in the algorithm, efficiency of the numerical
implementation, and accuracy required in
predicting the final endpoint of the therapy. However,
the results and discussion were focus on the validation
of the models. It was not obvious what is the trade-off
between the models?

{\color{red}
The authors agree. The focus of the manuscript was
the accuracy required in
predicting the final endpoint of the therapy. 
Remarks in the abstract on the trade-off between the
time investment in the algorithm, efficiency of the numerical
implementation, and accuracy required in
predicting the final endpoint of the therapy diffuse the focus
of the paper; these remarks have been (1) removed from the abstract and (2)
re-organized as results and  discussion points with model size and
computation time as you suggestion below.  Thank you for this suggestion.
}

\item What is accuracy required in predicting the final endpoint of the therapy?

{\color{red}

Regarding the topic of what accuracy is required to consider a model useful
for prediction, it is a multifaceted concept. Firstly, we consider a model
prediction as useful in a particular case if the models ablation region and
the ablation as indicated via MR temperature imaging has a Dice similarity
coefficient (DSC) greater than 0.7. Ultimately the quantitative value of
DSC=.7 is based on prior information and experiences in image processing
literature~\cite{yung2010quantitative,Dice1945measures,zou2004three};  
in which DSC$>$.7 is considered `good agreement'. This point is highlighted
in the discussion section where the DSC is introduced.
 Secondly, the models overall accuracy in
a cohort is measured by the ratio of useful predictions versus inaccurate
predictions (\textit{i.e.}, number of DSC $\geq$ 0.7 versus DSC $<$ 0.7).
This overall accuracy is desired to be as high as possible, but it is beyond the
scope of our present investigation to name a success rate that would
indicate the model would be useful in clinical practice. This manuscript
focuses on reporting the accuracy achieved for our current computational
methods. Future work will consider the accuracy needed to impact treatment 
on a very carefully selected patient population.

}
\item
It is not clear how this method will improve pretreatment planning?

{\color{red}
Long term project goals are to develop a data driven computer model
that can prospectively predict the thermal dose outcome with high
accuracy. These efforts present results towards these goals
in the form of a retrospective analysis
with sufficiently large number datasets to evaluate prediction
statistics within a hypothesis testing framework. 
Our LOOCV analysis attempts to mimic our targeted prospective scenario
in that the model is trained on all previous (N-1) datasets and
the trained model prediction accuracy is evaluated on the
independent remaining dataset.
The success or failure of this paradigm is
to be related to several factors. 1) The model used must be able to
describe clinically relevant ablations. 2) There must be a sufficient
quantity of retrospective datasets for the
training to converge. 3) The cohort of retrospective datasets must have sufficient similarity within the
group and to the prediction scenario. \textit{E.g.}, a model trained on datasets from liver ablations is unlikely to
be predictive for brain. However, a training cohort of brain
metastases may or may not have enough homogeneity for a trained model
to be predictive. These points are added at the beginning of
the Discussion section.
}
\item
The introduction is too long. It is not clear how the second part that
describes the hardware relates to the paper.  There is no information
on the model size and the actual computations time.
It is not clear which hardware was used with which particular model.

{\color{red}
The hardware discussion was intended to
motivate the discussion on the tradeoff between algorithm efficiency and
accuracy on state of the art computing architectures. 
Similar to the previous comment, this diffuses the modeling accuracy focus of the 
manuscript and has been reorganized to a discussion point.

The time required for computation is described in your
1\textsuperscript{st} critique and the Results and Discussion include that information now.

Thank you for returning our focus on the important
trade-off between accuracy and required time.
The accuracy of the models is reported, but the time of implementation
is not as you stated. The models implemented in the manuscript are 1) a homogenous,
time-dependent FEM model and 2) a homogeneous steady state model. The
FEM model is more accurate than the steady state model but takes considerably longer time
to run. It is difficult to succinctly express the difference in
computation times for at least two
reasons. During optimization for a given patient,
the FEM model and steady state model may have a different
number of optimization realizations. Secondly, the FEM model is transient and has multiple time steps within
one realization whereas the steady state model naturally has a single time step. Therefore, we
consider the simplest means to summarize computation time is to report the timing of optimization and LOOCV.
Optimization of the same dataset may involve differing number of realizations for the two models; LOOCV only
has one forward prediction for each dataset. In optimization, the FEM model takes {\color{green}[Some estimate of time]}; during LOOCV, it took
{\color{green}[Some estimate of time]}. The steady
state models LOOCV takes 246.6 seconds or 11.2 seconds on average per dataset. This indicates that both models
could be used before and during surgery. The Results
and Discussion have been amended to describe this
information; the new text is red.
}
\end{enumerate}

Reviewer 2:\\
Comments and suggestions to improve the manuscript:\\
\\
\begin{itemize}
\item
Please expand on any heating induced changes in model parameters. Was blood perfusion held constant for the
entire treatment duration? Blood perfusion is well
known to change during ablative exposures (as the
authors note, this is sometimes exploited to confirm post-ablation heating with contrast-enhanced imaging).  Have the authors explored any techniques for
incorporating temperature/thermal damage dependencies for blood perfusion? Suggest adjusting nominal
perfusion level based on time-temperature history (see, for example, Schutt, Med Phys, 2008).

@@

{\color{red}
The model parameters do not change with heating. We obviously must make this fact abundantly clear in the
manuscript and directly address the reasons why it is appropriate. Thank you for bringing to our attention
the lack of clarity regarding the use of model parameters. Accordingly, the manuscript has been
modified in the following places:
{\color{green}[list areas that were modified]}.

As the hyperthermia/ablation literature states in concert, tissue properties are dependent on temperature
and thermal damage state and these dependencies are important to consider in order to accurately predict
outcomes. However, this time-dependent phenomenon is mitigated in our investigation via shorter ablation
times relative to other thermal modalities. \textit{E.g.}, laser ablations typically last 1.5 minutes to 2 minutes
whereas the RFA procedures modeled in Schutt, 2008, were 12 and 15 minutes in duration. Furthermore, and
perhaps more importantly, we do not consider the parameters from model optimization (or model
training) to be truly physical. In this investigation, the modeling is driven by model training, not model
complexity/physical fidelity. 
}\\
\\
\item
Did the authors investigate the impact of temperature dependent optical properties? During the model
calibration procedure, it appears a single optical attenuation coefficient was determined for all spatial
location/time points. Since optical absorption properties likely change during the course of heating,
one approach may be to perform the model calibration over only the first ~30-60 seconds, to estimate the
nominal optical absorption coefficient. This nominal value could then modified during heating based on
previously published reports of the relationship between optical attenuation and temperature. Suggest
considering approach where

@@

{\color{red}
You are correct that the model calibration arrives at a constant effective optical attenuation coefficient,
eff, for one training dataset. Also, we certainly agree that patient-specific optimization via test pulse
or initial 30 seconds of ablation is an excellent idea for providing a calibrated parameter value. In this
manuscript we did not include patient-specific optimization because we did not think it would be
clinically practical. The optimization used in calibration requires [such and such] for the FEM model
and [such and such] for the steady state model. That optimization process could only begin after the low
power pulse the neurosurgeon uses for position confirmation. If the neurosurgeon is satisfied with the
localization, ablation would normally follow immediately. If a patient-specific optimization was
used, the clinician would have to wait for the optimization and then check the predicted ablation
zone.

Instead, our investigation is presented such that the present surgical technique would be mostly unmodified.
The surgical planning would be performed beforehand. During this planning, one prediction would take only
[such and such] for the FEM model and [such and such] for the steady state model. Even in the case that the
laser fiber was placed differently than the designed surgical plan, an updated prediction could be quick
generated. In summary, using parameter values calibrated by a prior cohort allows the surgically
planning to be performed within a tolerably short period of time. Furthermore, even in the future case
that computation was fast enough for patient-specific optimization, the surgeon would likely want a pre
surgical `best guess' to inform the surgical plan. This `best guess' is what is presented in this manuscript.
\\ \\
Our rationale for the use of constant parameters is explained in the answer to the previous critique you raised.
\\ \\
Finally, the last sentence in this critique seems to have been cut short. Please reiterate it, as necessity dictates.
}\\
\\
\item
Please state the test performed to confirm the discretization (mesh size, time steps) did not impact the solution.

{\color{red}
{\color{green}[DF, what did Rice CAAM do for this?]}
}\\
\\
\item
 Figure 1, consider blowing up the sketch of applicator (bottom half of fig 1a) so details can be distinguished more easily.

{\color{red}
The figure has been modified accordingly at
{\color{green}[page and line number]}. 
}\\
\\
\item
p6, L24 -- State units of frequency factor and activation energy

{\color{red}
The units have been added.
}\\
\\
\item
p6, L30 -- can the low-power pulse (used to confirm applicator positioning) be used to estimate optical
attenuation in a patient-specific manner? Did you consider evaluating this dataset and comparing to the
average value obtained from the calibration process for the other datasets?

@@

{\color{red}
We consider it a very good idea to use the low-power pulse as a patient-specific optimization dataset. Also,
your suggestion of comparing the optimized parameter values from the test pulses to the ablative pulses is
very astute. For now, patient-specific optimization is not close to being clinically feasible using the
methods of this manuscript.

Patient-specific optimization is mostly impeded by the fact that it stops the surgical workflow. In order to
acquire the low-power dataset, the neurosurgeon must first place the laser applicator. Then the test pulse
occurs and is imaged. That dataset is then pushed from the scanner to the computational architecture. There,
temperature maps must be generated. Only then can optimization begin, followed by a forward prediction
using the optimal optical parameter value. That forward prediction would then be used by the neurosurgeon for
surgical planning. We expect that the delay in the surgery incurred by patient-specific optimization would
preclude it being useful for some time. 
}\\
\\
\item
Please modify references so they appear in IJH style.

{\color{red}
The references are updated to match the \textit{IJH} style.
}\\
\\
\item
Figure 3, suggesting use same bin width for both histograms so the results for the two models can be
more readily compared.

{\color{red}
The figure has been amended.
}
\end{itemize}
\end{document}

### 2015-04-16 gpuPlanningAug13.tex
Fixed Table 1
WIP: paper updates not ready to merge

\documentclass[12pt]{article} 
\usepackage{amssymb,amsfonts,amsmath}
\usepackage{color,graphicx}
\usepackage{siunitx}
\sisetup{
per-mode=fraction,
fraction-function=\tfrac
}
\usepackage{fixltx2e}
\usepackage[nomarkers,tablesfirst]{endfloat}
\usepackage[left=0.7in,right=0.7in,top=0.7in,bottom=0.7in]{geometry}
\newcommand{\eqn}[1]{(\ref{#1})}
\newcommand{\numberofpatientsLOOCV}{22 }
\linespread{1.6}
%http://latex2rtf.sourceforge.net/latex2rtf_1_9_19.html#Conditional-Parsing
%Starting with LaTeX2RTF 1.9.18, there is a handy method for
%controlling which content should be processed by LaTeX or by
%LaTeX2RTF . Control is achieved using the standard \if facility of
%TeX. If you include the following line in the preamble of your document 
%
%     \newif\iflatextortf
%Then you will create a new \iflatextortf command in LaTeX . TeX
%sets the value of this to false by default. Now, LaTeX2RTF
%internally sets \iflatextortf to be true, and to ensure that this
%is always the case, LaTeX2RTF ignores the command
%\latextortffalse. This means that you can control how different
%applications process your document by
%
%     \iflatextortf
%     This code is processed only by latex2rtf
%     \else
%     This code is processed only by latex
%     \fi
%Note that \iflatextortf will only work within a section; you
%cannot use this command to conditionally parse code that crosses
%section boundaries. Also, it will only work on complete table or
%figure environments. Due to the mechanism used by LaTeX2RTF in
%processing these environments, at this time the only way to
%conditionally parse tables and figures is to include two complete
%versions of the environment in question, nested within an
%appropriate \iflatextortf structure.
%
\newif\iflatextortf

@@

\begin{abstract}
%Purpose
A cross validation analysis evaluating computer model predictions for \textit{a priori} 
planning magnetic resonance-guided laser induced thermal therapy
(MRgLITT) procedures in treating focal diseased brain tissue is presented.
Two mathematical models are selected to evaluate the
trade-off between the time investment in the algorithm, efficiency of the
numerical implementation, and accuracy required in predicting the final endpoint
of the therapy.  
%Methods
(1) A spectral element discretization of the transient Pennes bioheat transfer
equation implemented on GPU computing architectures predicts the
laser induced transient heating in perfused tissue. 
(2) A computationally inexpensive, closed-form algorithm for predicting the steady state heat transfer
from a linear superposition of analytic point source heating functions is also
considered.
Prediction accuracy is
retrospectively evaluated in MR thermometry data obtained during an
MRgLITT procedure in $N$ = \numberofpatientsLOOCV MR thermometry datasets via leave-one-out cross validation (LOOCV).
Modeling predictions are quantitatively evaluated
against MR thermal imaging in terms of a Dice similarity coefficient (DSC) between the
simulation and experimental data.
%Results
During LOOCV analysis, the transient model's DSC mean and median is 0.7323 and 0.8001, respectively,
with 15 of \numberofpatientsLOOCV DSC values exceeding the success criterion of DSC $\geq$ 0.7.
The steady state model's DSC mean and median is 0.6431 and 0.6770, respectively,
with 10 of \numberofpatientsLOOCV passing.
A one-sample, one-sided Wilcoxon signed rank test indicates the transient FEM model's
median DSC value is $\geq$ 0.7 at a statistically significant level. 
%The null hypothesis was accepted for the steady state model.
%Conclusions
Further, the spectral element discretization of the transient bioheat transfer model
provides greater potential in accurately calibrating the tissue properties needed
in treatment prediction.
%under an Ockham razor philosophy preferential to simplicity,
%the steady state superposition model with fewer floating
%point operations would be selected with respect to outcome prediction.
However, in the appropriate context, both approaches are likely to provide useful
information in guiding treatment decisions.

@@

Previous real-time implementations of the kernel for
predicting the laser induced bioheat transfer were made feasible via
implementation on a massively parallel supercomputers~\cite{fuentesetal09,
Odenetal07,Dilleretal07,Fengetal07,Fuentesetal08}. 
Despite the increasing presence of large computational clusters within
large academic hospitals, reliance on such an approach makes clinical
translation of this research extremely difficult. This technology must
be practically available to any clinic using an MR-guided thermal therapy
system whereas integration with an institution's cluster and intranet creates
significant overhead and potential complications.
While modern single-chip multi-core CPU's have a peak performance of
$\mathcal{O}(10-100)$ GFLOPs and are sufficient 
for inexpensive algorithms such as the linear superposition of analytic functions presented
in Section~\ref{AnalyticBioheatModel}, 
current GPUs~\cite{Farber2011a,Muller2013} have a peak
performance of $\mathcal{O}(1-2)$ TFLOPs (single precision) and the
portability of these devices are ideal for developing more advanced 
numerical discretizations of the underlying partial differential equations
for guiding the therapy. However, a significant
time investment is needed to express the class of memory bound numerical
methods used in simulation within an algorithmic framework that can
simultaneously exploit these floating point capabilities and minimize the
global memory access latency of these devices.
% 28 = 1030/ (148/4) single precision.
% max floating point operations per second is 1030 GFlop
% data transfer of single precision numbers to the cores = bandwidth x data type size = 148  x (4bytes / float)
% >>> 1030./ (148./4.)
% 27.83783783783784
For example, the $\mathcal{O} (150)$~GB/s memory bandwidth of the Tesla M2070
used in the manuscript requires $\approx$ 30 floating point operations per
single precision global memory access to fully utilize the floating point capabilities
of the GPU. 
The spectral element method presented in Section~\ref{BioheatComputerModel}
achieves efficient global memory access through a matrix-free
implementation of the inherent linear system of equations.
As opposed to explicitly storing and reading matrix entries
from global memory, the matrix-free method recalculates the local matrix entries
needed within the linear system solve.
This memory access design pattern has demonstrated a 4-10x speedup  over
the matrix explicit methods~\cite{Muller2013,Knepley_2_2013,Medina2014}.

@@

\begin{tabular}{|c|c|c|c|c|c|} \hline 
\textit{k} [$\frac{\si{\watt}}{\si{\metre\kelvin}}$]   & $\omega$ [$\frac{\si{\kilo\gram}}{\si{\metre\cubed\second}}$] &  \textit{g} [Unity]  &  $\mu_s$ [\si{\per\centi\metre}] &   $\mu_a$ [\si{\per\centi\metre}] & $\rho$ [$\frac{\si{\kilo\gram}}{\si{\metre\cubed}}$] \\ \hline
          0.527              &             9.0             & 0.88  &       14000              &       500             &  1045    \\ \hline
$c_{blood}$ [$\frac{\si{\joule}}{\si{\kilo\gram\kelvin}}$] &  $c_{p}$ [$\frac{\si{\joule}}{\si{\kilo\gram\kelvin}}$]& \textit{h} [$\frac{\si{\watt}}{\si{\kelvin\metre\cubed}}$] & $u_{cooling}$ [\si{\degreeCelsius}]& $u_a$ [\si{\degreeCelsius}] \\ \hline
     3840                      &                  3600  & 100 & 21   & 37 \\ \hline
\end{tabular}
%\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|c|} \hline 
%\textit{k} [$\frac{\si{\watt}}{\si{\metre\kelvin}}$]   & $\omega$ [$\frac{\si{\kilo\gram}}{\si{\metre\cubed\second}}$] &  \textit{g} [Unity]  &  $\mu_s$ [\si{\per\centi\metre}] &   $\mu_a$ [\si{\per\centi\metre}] & $\rho$ [$\frac{\si{\kilo\gram}}{\si{\metre\cubed}}$] & $c_{blood}$ [$\frac{\si{\joule}}{\si{\kilo\gram\kelvin}}$] &  $c_{p}$ [$\frac{\si{\joule}}{\si{\kilo\gram\kelvin}}$]& \textit{h} [$\frac{\si{\watt}}{\si{\kelvin\metre\cubed}}$] & $u_{cooling}$ [\si{\degreeCelsius}]& $u_a$ [\si{\degreeCelsius}]\\ \hline
%          0.527              &             9.0             & 0.88  &       14000              &       500                  &  1045                        &            3840                      &                  3600  & 100 & 21   & 37    \\ \hline
%\end{tabular}
\end{table*}
\fi
%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

@@

The trained model's predictive ability is
evaluated by analyzing the distribution of DSC values from the $N$ iterations of
LOOCV. 
A one-sample, one-sided Wilcoxon signed rank test examines if the trained
model prediction's median exceeds DSC $\geq$ 0.7. 
One-sample calculations were computed using a threshold DSC value, $0.7$, as 
the null hypothesis, $H_0$.
\[
   H_0: DSC  = 0.7  \qquad   H_1: DSC > 0.7
\]
The value chosen is a commonly accepted value in image processing
literature~\cite{yung2010quantitative,Dice1945measures,zou2004three};  
$DSC=1$ implies complete agreement between the measure and predicted damage model,
$i.e.$ the predicted and measured damages volumes completely overlap.
A two-sample, paired Wilcoxon signed rank test was used to compare if the two models'
prediction medians are statistically different from one another.
All statistical tests and descriptive statistics were evaluated on
GraphPad Prism 6.01 (GraphPad Software, La Jolla, CA).

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

>>>>

\documentclass[12pt]{article} 
\usepackage{amssymb,amsfonts,amsmath}
\usepackage{color,graphicx}
\usepackage{siunitx}
%\sisetup{
%per-mode=fraction,
%fraction-function=\tfrac
%}
\usepackage{fixltx2e}
\usepackage[nomarkers,tablesfirst]{endfloat}
\usepackage[left=0.7in,right=0.7in,top=0.7in,bottom=0.7in]{geometry}
\newcommand{\eqn}[1]{(\ref{#1})}
\newcommand{\numberofpatientsLOOCV}{22 }
\linespread{1.6}
%http://latex2rtf.sourceforge.net/latex2rtf_1_9_19.html#Conditional-Parsing
%Starting with LaTeX2RTF 1.9.18, there is a handy method for
%controlling which content should be processed by LaTeX or by
%LaTeX2RTF . Control is achieved using the standard \if facility of
%TeX. If you include the following line in the preamble of your document 
%
%     \newif\iflatextortf
%Then you will create a new \iflatextortf command in LaTeX . TeX
%sets the value of this to false by default. Now, LaTeX2RTF
%internally sets \iflatextortf to be true, and to ensure that this
%is always the case, LaTeX2RTF ignores the command
%\latextortffalse. This means that you can control how different
%applications process your document by
%
%     \iflatextortf
%     This code is processed only by latex2rtf
%     \else
%     This code is processed only by latex
%     \fi
%Note that \iflatextortf will only work within a section; you
%cannot use this command to conditionally parse code that crosses
%section boundaries. Also, it will only work on complete table or
%figure environments. Due to the mechanism used by LaTeX2RTF in
%processing these environments, at this time the only way to
%conditionally parse tables and figures is to include two complete
%versions of the environment in question, nested within an
%appropriate \iflatextortf structure.
%
\newif\iflatextortf

@@

\begin{abstract}
%Purpose
{\color{red}
A cross validation analysis evaluating computer model prediction accuracy
for \textit{a priori} planning magnetic resonance-guided laser induced
thermal therapy (MRgLITT) procedures in treating focal diseased brain
tissue is presented.
Two mathematical models are considered.
%Two mathematical models are selected to evaluate the
%trade-off between the time investment in the algorithm, efficiency of the
%numerical implementation, and accuracy required in predicting the final endpoint
%of the therapy.  
}
%Methods
(1) A spectral element discretization of the transient Pennes bioheat
transfer equation is implemented to 
%implemented on GPU computing architectures 
predict the laser induced heating in perfused tissue. 
(2) A 
%computationally inexpensive, 
closed-form algorithm for predicting the steady state heat transfer from a
linear superposition of analytic point source heating functions is also
considered.
{\color{red}
Prediction accuracy is
retrospectively evaluated 
via leave-one-out cross validation (LOOCV).  
Modeling predictions are quantitatively evaluated
in terms of a Dice similarity coefficient (DSC) between the
simulated thermal dose and
thermal dose information contained within
$N$ = \numberofpatientsLOOCV 
MR thermometry datasets.
%obtained during MRgLITT procedures. 
}
%Results
During LOOCV analysis, the transient model's DSC mean and median is 0.7323 and 0.8001, respectively,
with 15 of \numberofpatientsLOOCV DSC values exceeding the success criterion of DSC $\geq$ 0.7.
The steady state model's DSC mean and median is 0.6431 and 0.6770, respectively,
with 10 of \numberofpatientsLOOCV passing.
A one-sample, one-sided Wilcoxon signed rank test indicates that the
transient FEM model achieves the prediction success critera, 
DSC $\geq$ 0.7, at a statistically significant level. 
%The null hypothesis was accepted for the steady state model.
%Conclusions
%Further, the spectral element discretization of the transient bioheat transfer model
%provides greater potential in accurately calibrating the tissue properties needed
%in treatment prediction.
%under an Ockham razor philosophy preferential to simplicity,
%the steady state superposition model with fewer floating
%point operations would be selected with respect to outcome prediction.
%However, in the appropriate context, both approaches are likely to provide useful
%information in guiding treatment decisions.

@@

@@

\begin{tabular}{|c|c|c|c|c|c|} \hline 
\textit{k} [$\frac{\si{\watt}}{\si{\metre\kelvin}}$]   & $\omega$ [$\frac{\si{\kilo\gram}}{\si{\metre\cubed\second}}$] &  \textit{g} [Unity]  &  $\mu_s$ [\si{\per\centi\metre}] &   $\mu_a$ [\si{\per\centi\metre}] & $\rho$ [$\frac{\si{\kilo\gram}}{\si{\metre\cubed}}$] \\ \hline
          0.527              &             9.0             & 0.88  &       14000              &       500             &  1045    \\ \hline
$c_{blood}$ [$\frac{\si{\joule}}{\si{\kilo\gram\kelvin}}$] &  $c_{p}$ [$\frac{\si{\joule}}{\si{\kilo\gram\kelvin}}$]& \textit{h} [$\frac{\si{\watt}}{\si{\kelvin\metre\cubed}}$] & $u_{cooling}$ [\si{\degreeCelsius}]& $u_a$ [\si{\degreeCelsius}] & \\ \hline
     3840                      &                  3600  & 100 & 21   & 37 & \\ \hline
\end{tabular}
%\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|c|} \hline 
%\textit{k} [$\frac{\si{\watt}}{\si{\metre\kelvin}}$]   & $\omega$ [$\frac{\si{\kilo\gram}}{\si{\metre\cubed\second}}$] &  \textit{g} [Unity]  &  $\mu_s$ [\si{\per\centi\metre}] &   $\mu_a$ [\si{\per\centi\metre}] & $\rho$ [$\frac{\si{\kilo\gram}}{\si{\metre\cubed}}$] & $c_{blood}$ [$\frac{\si{\joule}}{\si{\kilo\gram\kelvin}}$] &  $c_{p}$ [$\frac{\si{\joule}}{\si{\kilo\gram\kelvin}}$]& \textit{h} [$\frac{\si{\watt}}{\si{\kelvin\metre\cubed}}$] & $u_{cooling}$ [\si{\degreeCelsius}]& $u_a$ [\si{\degreeCelsius}]\\ \hline
%          0.527              &             9.0             & 0.88  &       14000              &       500                  &  1045                        &            3840                      &                  3600  & 100 & 21   & 37    \\ \hline
%\end{tabular}
\end{table*}
\fi
%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

@@

The trained model's predictive ability is
evaluated by analyzing the distribution of DSC values from the $N$ iterations of
LOOCV. 
A one-sample, one-sided Wilcoxon signed rank test examines if the trained
model prediction's median exceeds DSC $\geq$ 0.7. 
One-sample calculations were computed using a threshold DSC value, $0.7$, as 
the null hypothesis, $H_0$.
\[
   H_0: DSC  = 0.7  \qquad   H_1: DSC > 0.7
\]
{\color{red}
The value chosen is a commonly accepted value in image processing
literature~\cite{yung2010quantitative,Dice1945measures,zou2004three};  
$DSC=1$ implies complete agreement between the measure and predicted damage model,
$i.e.$ the predicted and measured damages volumes completely overlap.
}
A two-sample, paired Wilcoxon signed rank test was used to compare if the two models'
prediction medians are statistically different from one another.
All statistical tests and descriptive statistics were evaluated on
GraphPad Prism 6.01 (GraphPad Software, La Jolla, CA).

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
{\color{blue}
Previous real-time implementations of the kernel for
predicting the laser induced bioheat transfer were made feasible via
implementation on a massively parallel supercomputers~\cite{fuentesetal09,
Odenetal07,Dilleretal07,Fengetal07,Fuentesetal08}. 
Despite the increasing presence of large computational clusters within
large academic hospitals, reliance on such an approach makes clinical
translation of this research extremely difficult. This technology must
be practically available to any clinic using an MR-guided thermal therapy
system whereas integration with an institution's cluster and intranet creates
significant overhead and potential complications.
While modern single-chip multi-core CPU's have a peak performance of
$\mathcal{O}(10-100)$ GFLOPs and are sufficient 
for inexpensive algorithms such as the linear superposition of analytic functions presented
in Section~\ref{AnalyticBioheatModel}, 
current GPUs~\cite{Farber2011a,Muller2013} have a peak
performance of $\mathcal{O}(1-2)$ TFLOPs (single precision) and the
portability of these devices are ideal for developing more advanced 
numerical discretizations of the underlying partial differential equations
for guiding the therapy. However, a significant
time investment is needed to express the class of memory bound numerical
methods used in simulation within an algorithmic framework that can
simultaneously exploit these floating point capabilities and minimize the
global memory access latency of these devices.
% 28 = 1030/ (148/4) single precision.
% max floating point operations per second is 1030 GFlop
% data transfer of single precision numbers to the cores = bandwidth x data type size = 148  x (4bytes / float)
% >>> 1030./ (148./4.)
% 27.83783783783784
For example, the $\mathcal{O} (150)$~GB/s memory bandwidth of the Tesla M2070
used in the manuscript requires $\approx$ 30 floating point operations per
single precision global memory access to fully utilize the floating point capabilities
of the GPU. 
The spectral element method presented in Section~\ref{BioheatComputerModel}
achieves efficient global memory access through a matrix-free
implementation of the inherent linear system of equations.
As opposed to explicitly storing and reading matrix entries
from global memory, the matrix-free method recalculates the local matrix entries
needed within the linear system solve.
This memory access design pattern has demonstrated a 4-10x speedup  over
the matrix explicit methods~\cite{Muller2013,Knepley_2_2013,Medina2014}.
}
